Binning Strategies and Related Loss for Large Data

### 2015-04-13 Paper.Rnw
Frequency loss conclusions

We have presented two algorithms for spatially binning data points; the standard algorithm, which has rounding artifacts but generally lower loss, and the random algorithm, which is a fast version of linear binning and is somewhat more computationally efficient. The standard algorithm is superior in terms of net spatial loss, but requires more care in selecting binning parameters than random binning to avoid problematic visual features such as binning caused stripping. In setting binning parameters for each spatial dimension, binwidth should be an integer multiple of the data resolution and the origin should be offset half the data resolution below the lowest observed value.

>>>>

We have presented two algorithms for spatially binning data points; the standard algorithm, which has rounding artifacts but generally lower loss, and the random algorithm, which is a fast version of linear binning and is somewhat more computationally efficient. The standard algorithm is superior in terms of net spatial loss, but requires more care in selecting binning parameters than random binning to avoid generating problematic visual features. In setting binning parameters for each spatial dimension, binwidth should be an integer multiple of the data resolution and the origin should be offset half the data resolution below the lowest observed value.

We also have explored the loss of frequency information for cases where frequencies are also binned to create a discrete color scale for tile shading. It is recommended to use less than ten seperate shades due to bounds of perceptual discrimination but at least four shades to avoid the steep frequency losses seen with fewer frequency bins. Using quantile frequency binning and log transforming the bin counts prior to standard frequency binning are shown to be reasonable methods -- with slight different interpretability -- for handling situations with heavily skewed bin count distributions.
